\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-12-11 forwards-and-backwards.tex
.

We shall next show that $\zeta^\epsilon * \psi_t^\epsilon$ solves \eqref{PME1}
up to a remainder of order $\epsilon$.
First observe that
\begin{equation}
\label{total equation}
\int\Delta\left((\rho^{\epsilon}*\psi_t^\epsilon)\psi_t^\epsilon\right)\phi  dx
    =
    - \int \nabla(\rho^{\epsilon}*\psi_t^\epsilon)\, \psi_t^\epsilon\,\nabla\phi dx
    -\int \rho^{\epsilon}*\psi_t^\epsilon \,\nabla \psi_t^\epsilon \, \nabla\phi dx.
\end{equation}
We would like to show that this is close to $\int (\zeta^\epsilon * \psi^\epsilon_t)^2 \Delta \phi dx$.
For the first term
\begin{eqnarray}
\nonumber
\int \nabla(\rho^{\epsilon}*\psi_t^\epsilon)\, \psi_t^\epsilon\, \nabla\phi  dx
&=&
\int\int\int\nabla \psi_t^\epsilon(x-y)\zeta^{\epsilon}(y-z)
\check{\zeta}^{\epsilon}(z)
\psi_t^\epsilon(x)\nabla\phi(x) dz  dy  dx
\\
\nonumber
&=&
\int\int\int\nabla \psi_t^\epsilon(\tilde{x}-\tilde{y})
\zeta^{\epsilon}(\tilde{y})
\zeta^{\epsilon}(\tilde{z})\psi_t^\epsilon(\tilde{x}-\tilde{z})
\nabla\phi(\tilde{x}-\tilde{z}) d\tilde{z} d\tilde{y} d\tilde{x}
\\
\nonumber
&=&\int(\nabla \zeta^{\epsilon}*\psi_t^\epsilon)\,
\left(\zeta^{\epsilon}*(\psi_t^\epsilon\nabla\phi)\right)
d x\\
\nonumber
&=&
\frac{1}{2}\int\nabla ((\zeta^{\epsilon}*\psi_t^\epsilon)^2)
\,\nabla\phi  dx
\\
\label{error 1}
&&+
\int\nabla(\zeta^{\epsilon}*\psi_t^\epsilon)
\left[\zeta^{\epsilon}*(\psi_t^\epsilon\,\nabla\phi)-
\nabla\phi\,(\zeta^{\epsilon}*\psi_t^\epsilon)\right] dx,
\end{eqnarray}
where, as before, we have substituted $\tilde{x}=x-z$, $\tilde{y}=y-z$,
$\tilde{z}=-z$.
We are going to bound the square of the $L^2$-norm of the term in square brackets 
in~(\ref{error 1}) by the product of its $L^1$-norm (which is bounded
by a constant times the $L^1$-norm of $\psi_t^\epsilon$) and its $L^\infty$-norm.
To control the $L^\infty$-norm,
we use the Intermediate Value Theorem to see that
\begin{multline*}
\left|\int\left[\psi_t^\epsilon(x-y)\nabla\phi(x-y)\zeta^{\epsilon}(y)-
\nabla\phi(x)\psi_t^\epsilon(x-y)\zeta^{\epsilon}(y)
\right] dy\right|
\\
	\leq C\|\Delta\phi\|_\infty\int \psi_t^\epsilon (x-y)
	\epsilon\frac{|y|}{\epsilon}\zeta^{\epsilon}(y) d y
	\leq C\|\Delta\phi\|_\infty\epsilon \|z\zeta(z)\|_\infty\int \psi_t^\epsilon (x-y) dy.
	%|y|\zeta^{\epsilon}(y) d y
\end{multline*}
Since $\zeta\in\mathcal{S}(\IR)$, and as we checked above $\psi_t^\epsilon$ is 
uniformly bounded in $L^1$ over $[0,T]$,
this expression is
$\mathcal{O}(\epsilon)$.  \revpoint{1}{31}

>>>>

We shall next show that $\zeta^\epsilon * \psi_t^\epsilon$ solves \eqref{PME1}
up to a remainder of order $\epsilon$.
First observe that
\begin{equation}
\label{total equation}
\int\Delta\left((\rho^{\epsilon}*\psi_t^\epsilon)\psi_t^\epsilon\right)\phi  dx
    =
    - \int \nabla(\rho^{\epsilon}*\psi_t^\epsilon)\, \psi_t^\epsilon\,\nabla\phi dx
    -\int \rho^{\epsilon}*\psi_t^\epsilon \,\nabla \psi_t^\epsilon \, \nabla\phi dx.
\end{equation}
We would like to show that this is close to $\int (\zeta^\epsilon * \psi^\epsilon_t)^2 \Delta \phi dx$.
For the first term
\begin{eqnarray}
\nonumber
\int \nabla(\rho^{\epsilon}*\psi_t^\epsilon)\, \psi_t^\epsilon\, \nabla\phi  dx
&=&
\int\int\int\nabla \psi_t^\epsilon(x-y)\zeta^{\epsilon}(y-z)
\check{\zeta}^{\epsilon}(z)
\psi_t^\epsilon(x)\nabla\phi(x) dz  dy  dx
\\
\nonumber
&=&
\int\int\int\nabla \psi_t^\epsilon(\tilde{x}-\tilde{y})
\zeta^{\epsilon}(\tilde{y})
\zeta^{\epsilon}(\tilde{z})\psi_t^\epsilon(\tilde{x}-\tilde{z})
\nabla\phi(\tilde{x}-\tilde{z}) d\tilde{z} d\tilde{y} d\tilde{x}
\\
\nonumber
&=&\int(\nabla \zeta^{\epsilon}*\psi_t^\epsilon)\,
\left(\zeta^{\epsilon}*(\psi_t^\epsilon\nabla\phi)\right)
d x\\
\nonumber
&=&
\frac{1}{2}\int\nabla ((\zeta^{\epsilon}*\psi_t^\epsilon)^2)
\,\nabla\phi  dx
\\
\label{error 1}
&&+
\int\nabla(\zeta^{\epsilon}*\psi_t^\epsilon)
\left[\zeta^{\epsilon}*(\psi_t^\epsilon\,\nabla\phi)-
\nabla\phi\,(\zeta^{\epsilon}*\psi_t^\epsilon)\right] dx,
\end{eqnarray}
where, as before, we have substituted $\tilde{x}=x-z$, $\tilde{y}=y-z$,
$\tilde{z}=-z$.
We are going to bound the square of the $L^2$-norm of the term in square brackets 
in~(\ref{error 1}) by the product of its $L^1$-norm (which is bounded
by a constant times the $L^1$-norm of $\psi_t^\epsilon$) and its $L^\infty$-norm.
To control the $L^\infty$-norm,
we use the Intermediate Value Theorem to see that
\begin{multline*}
\left|\int\left[\psi_t^\epsilon(x-y)\nabla\phi(x-y)\zeta^{\epsilon}(y)-
\nabla\phi(x)\psi_t^\epsilon(x-y)\zeta^{\epsilon}(y)
\right] dy\right|
\\
	\leq C\|\Delta\phi\|_\infty\int \psi_t^\epsilon (x-y)
	\epsilon\frac{|y|}{\epsilon}\zeta^{\epsilon}(y) d y
	\leq C\|\Delta\phi\|_\infty\epsilon \|z\zeta(z)\|_\infty\int \psi_t^\epsilon (x-y) dy.
	%|y|\zeta^{\epsilon}(y) d y
\end{multline*}
Since $\zeta\in\mathcal{S}(\IR)$,
and (as we checked above)
$\psi_t^\epsilon$ is uniformly bounded in $L^1$ over $[0,T]$,
this expression is $\mathcal{O}(\epsilon)$ -- \revpoint{1}{31}
i.e., is bounded by a constant multiple of $\epsilon$
with a constant that does not depend on $t \in [0,T]$ or $x$.
